Data Specialization for a Staged Lambda Calculus

### 2015-02-27 paper/splitting.tex
fixed yhat

\paragraph{Tuple.} For an example of splitting terms with more than one argument, consider the tuple $\tup{e_1,e_2}$,
where $\diaone {e_k} {\xi_k} {v_k}$ and $\rtab {\xi_k} {v_k} \vsplito
\mval {i_k} {q_k}$ for $k = 1,2$.  Inductively:
\[
	\splitone {e_k} A {c_k} {l_k.r_k} \text{ and } \reduce {c_k} {\tup{i_k,b_k}} \text{ and } [b_k/l_k]r_k \equiv q_k
\]
Since $\diaone {\tup{e_i,e_2}} {\xi_1,\xi_2} {\tup{v_1,v_2}}$ 
and $\rtab {\xi_1,\xi_2} {\tup{v_1,v_2}} \vsplito \mval {\tup{i_1,i_2}} {\tup{q_1,q_2}}$,
splitting generates:
\[
	\reduce {c} {\tup{\tup{i_1,i_2},b}} \text{ and } [b/l]r \equiv \tup{q_1,q_2}
\]
The resulting boundary value is the tuple of the boundary values of
the subterms, $b = \tup{b_1,b_2}$.  The value is 
constructed using,
\[
c = \letin{\tup{y_1,z_1}}{c_1}{\letin{\tup{y_2,z_2}}{c_2}{\tup{\tup{y_1, y_2},\tup{z_1, z_2}}}}
\]
And deconstructed with a tuple pattern, $l.r = \tup{l_1,l_2}.\tup{r_1,r_2}$,

\paragraph {Case}
Splitting is most complicated at $\caseof {e_1} {x_2.e_2} {x_3.e_3}$ terms because of their divergent control flow.
Consider that when $\reduce {e_1} {\inl {v_1}}$, both the first and second stage component of $e_2$ will be executed, 
but $e_3$ will be ignored entirely. The reverse occurs when $\reduce {e_1} {\inr {v_1}}$.
This implies that both the first and second stage results of splitting will need to have \texttt{case} statements.

@@

\paragraph{Next and Prev}
Since \texttt{next} and \texttt{prev} do nothing more than convert between 
second stage terms at \bbtwo\ and second stage terms encapsulated within $\fut$s at \bbonem,
their splitting rules are relatively simple.

Given an $\splittwo e ? p l r$ and $\diatwo e q$ for which the splitting invariant holds, we know
\[
	\reduce p b \text{ and } [b/l]r \equiv q
\]
The splitting invariant for $\splitone {\next {e'}} ? {c'} {l'.r'}$ 
and $\diaone {\next {e'}} {\yhat \mapsto q} {\next \yhat}$
and $\rtab {\yhat \mapsto q} {\next \yhat} \vsplito \mval {\tup{}} {\letin \yhat q \yhat}$
is
\[
	\reduce {c'} {\tup {\tup{},b'}} \text{ and } [b'/l']r' \equiv \letin \yhat q \yhat
\]
which can be fulfilled using $c' = \tup{\tup{},p}$ and $l.r = l'.r'$.
Applying this process in reverse yields a splitting rule for $\prev e$ which applies \texttt{pi1} to the combined term of $e$.

>>>>

\paragraph{Tuple.} For an example of splitting terms with more than one argument, consider the tuple $\tup{e_1,e_2}$,
where $\diaone {e_k} {\xi_k} {v_k}$ and 
$\rtab {\xi_k} {v_k} \vsplito \mval {i_k} {q_k}$ for $k = 1,2$.  
Inductively:
\[
	\splitone {e_k} A {c_k} {l_k.r_k} \text{ and } \reduce {c_k} {\tup{i_k,b_k}} \text{ and } [b_k/l_k]r_k \equiv q_k
\]
Since $\diaone {\tup{e_i,e_2}} {\xi_1,\xi_2} {\tup{v_1,v_2}}$ 
and $\rtab {\xi_1,\xi_2} {\tup{v_1,v_2}}$ and $\tup{v_1,v_2} \vsplito \mval {\tup{i_1,i_2}} {\tup{q_1,q_2}}$,
splitting generates:
\[
	\reduce {c} {\tup{\tup{i_1,i_2},b}} \text{ and } [b/l]r \equiv \tup{q_1,q_2}
\]
The resulting boundary value is the tuple of the boundary values of
the subterms, $b = \tup{b_1,b_2}$.  The value is 
constructed using,
\[
c = \letin{\tup{y_1,z_1}}{c_1}{\letin{\tup{y_2,z_2}}{c_2}{\tup{\tup{y_1, y_2},\tup{z_1, z_2}}}}
\]
And deconstructed with a tuple pattern, $l.r = \tup{l_1,l_2}.\tup{r_1,r_2}$,

\paragraph {Case}
Splitting is most complicated at $\caseof {e_1} {x_2.e_2} {x_3.e_3}$ terms because of their divergent control flow.

We consider the two possible control paths seperately.  In the first case, 
$\diaone {e_1} {\xi_1} {\inl {v_1}}$ and $\rtab {\xi_1} {\inl {v_1}} \vsplito \mval {\inl {i_1}} {q_1}$.
Inductively,
\[
	\splitone {e_1} A {c_1} {l_1.r_1} \text{ and } \reduce {c_1} {\tup{\inl {i_1},b_1}} \text{ and } [b_1/l_1]r_1 \equiv q_1
\]
Additionally, the branch first evaluates $\diaone {[v_1/x_2]e_2} {\xi_2} {v_2}$ and $\rtab {\xi_2} {v_2} \vsplito \mval {i_2} {q_2}$.
Inductively,
\[
	\splitone {e_2} A {c_2} {l_2.r_2} \text{ and } \reduce {[i_1/x_2]e_2} {\tup{i_2,b_2}} \text{ and } [b_2/l_2](\letin x {q_1} {r_2}) \equiv q_2
\]

So we need to satisfy.....

Consider that when $\reduce {e_1} {\inl {v_1}}$, both the first and second stage component of $e_2$ will be executed, 
but $e_3$ will be ignored entirely. The reverse occurs when $\reduce {e_1} {\inr {v_1}}$.
This implies that both the first and second stage results of splitting will need to have \texttt{case} statements.

@@

\paragraph{Next and Prev}
Since \texttt{next} and \texttt{prev} do nothing more than convert between 
second stage terms at \bbtwo\ and second stage terms encapsulated within $\fut$s at \bbonem,
their splitting rules are relatively simple.
Given an $\splittwo e ? p l r$ and $\diatwo e q$, the splitting invariant says,
\[
	\reduce p b \text{ and } [b/l]r \equiv q
\]
The splitting invariant for $\splitone {\next {e'}} ? {c'} {l'.r'}$,
where $\diaone {\next {e'}} {{\hat y} \mapsto q} {\next {\hat y}}$
and $\rtab {{\hat y} \mapsto q} {\next {\hat y}} \vsplito \mval {\tup{}} {\letin {\hat y} q {\hat y}}$
then requires
\[
	\reduce {c'} {\tup {\tup{},b'}} \text{ and } [b'/l']r' \equiv \letin {\hat y} q {\hat y}
\]
which can be fulfilled using $c' = \tup{\tup{},p}$ and $l'.r' = l.r$.
Applying this process in reverse yields a splitting rule for $\prev e$ which applies \texttt{pi1} to the combined term of $e$.
